### 2025-04-09 analysis/primary_analyses.qmd
work on including tables; update stuff

---
title: "Primary analyses"
---

## Load packages

```{r}
library(foreign, quietly = TRUE, warn.conflicts = FALSE)
library(tidyverse, quietly = TRUE, warn.conflicts = FALSE)
library(semTools, quietly = TRUE, warn.conflicts = FALSE)
library(lavaan, quietly = TRUE, warn.conflicts = FALSE)
library(MBESS, quietly = TRUE, warn.conflicts = FALSE)
library(corx)
```

## Load datasets

```{r}
dat_poly <- read.spss("../data/Haupterhebung Studie III (Polytom 6PS).sav", 
  to.data.frame = TRUE, use.value.labels = FALSE, na.omit = FALSE)
dat_dich <- read.spss("../data/Haupterhebung Studie III (Dichotom 2PS).sav",
  to.data.frame = TRUE, use.value.labels = FALSE, na.omit = FALSE)
```

## Fit and analyze CFA models

### Tau-congeneric model for each trait

```{r}
### tau-congeneric model without modification 
modelAP <- "
  f1 =~ A_1 +  A_2 + A_3 + A_4 + A_5 + A_6 + A_7 + A_8 + A_9 +A_10 + A_11 + A_12 + Alter_realW"
modelGP <- "
  f1 =~ KGr_13 + KGr_14 + KGr_15 + KGr_16 + KGr_17 + KGr_18 + KGr_19 + KGr_20 + KGr_21 + KGr_22 + KGr_23 + KGr_24 + Koerpergroesse_realW"
modelGeP <- "
  f1 =~ KGe_25 + KGe_26 + KGe_27 + KGe_28 + KGe_29 + KGe_30 + KGe_31 + KGe_32 + KGe_33 + KGe_34 + KGe_35 + KGe_36 + Koerpergewicht_realW"

fit1A6P <- cfa(model = modelAP, data = dat_poly, ordered = c("A_1","A_2", "A_3", "A_4", "A_5", "A_6", "A_7","A_8", "A_9", "A_10", "A_11","A_12"), std.lv=TRUE, meanstructure=TRUE)
fit1A2P <- cfa(model = modelAP, data = dat_dich, ordered = c("A_1","A_2", "A_3", "A_4", "A_5", "A_6", "A_7","A_8", "A_9", "A_10", "A_11","A_12"), std.lv=TRUE, meanstructure=TRUE)
fit1G6P <- cfa(model = modelGP, data = dat_poly, ordered = c("KGr_13","KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19","KGr_20", "KGr_21", "KGr_22", "KGr_23","KGr_24"), std.lv=TRUE, meanstructure=TRUE)
fit1G2P <- cfa(model = modelGP, data = dat_dich, ordered = c("KGr_13","KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19","KGr_20", "KGr_21", "KGr_22", "KGr_23","KGr_24"), std.lv=TRUE, meanstructure=TRUE)
fit1Ge6P <- cfa(model = modelGeP, data = dat_poly, ordered = c("KGe_25","KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31","KGe_32", "KGe_33", "KGe_34", "KGe_35","KGe_36"), std.lv=TRUE, meanstructure=TRUE)
fit1Ge2P <- cfa(model = modelGeP, data = dat_dich, ordered = c("KGe_25","KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31","KGe_32", "KGe_33", "KGe_34", "KGe_35","KGe_36"), std.lv=TRUE, meanstructure=TRUE)

summary(fit1A6P, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1A2P, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1G6P, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1G2P, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1Ge6P, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1Ge2P, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)

tabelle2a <- compareFit(fit1A6P, fit1G6P, fit1Ge6P, nested=FALSE)
tabelle2b <- compareFit(fit1A2P, fit1G2P, fit1Ge2P, nested=FALSE)
summary(tabelle2a)
summary(tabelle2b)
```

```{r}
# Funktion zum Extrahieren standardisierter Ladungen (nur einmal definieren)
get_std_loadings <- function(fit, model_name) {
  standardizedSolution(fit) %>%
    filter(op == "=~") %>%
    select(lhs, rhs, est.std) %>%
    mutate(model = model_name)
}

# ---- Alter ----
loadings1 <- get_std_loadings(fit1A6P, "age 1")
loadings2 <- get_std_loadings(fit1A2P, "age d1")
all_loadings12 <- bind_rows(loadings1, loadings2)

table_loadings12 <- all_loadings12 %>%
  pivot_wider(names_from = model, values_from = est.std)

print(table_loadings12)

# ---- Größe ----
loadings3 <- get_std_loadings(fit1G6P, "height 2")
loadings4 <- get_std_loadings(fit1G2P, "height d2")
all_loadings34 <- bind_rows(loadings3, loadings4)

table_loadings34 <- all_loadings34 %>%
  pivot_wider(names_from = model, values_from = est.std)

print(table_loadings34)

# ---- Gewicht ----
loadings5 <- get_std_loadings(fit1Ge6P, "weight 3")
loadings6 <- get_std_loadings(fit1Ge2P, "weight d3")
all_loadings56 <- bind_rows(loadings5, loadings6)

table_loadings56 <- all_loadings56 %>%
  pivot_wider(names_from = model, values_from = est.std)

print(table_loadings56)
```

### Bifactor models

```{r}
### Residual Psychologie
modelA_p <- "
  age =~ A_1 + A_2 + A_3 + A_4 + A_5 + A_6 + A_7 + A_8 + A_9 + A_10 + A_11 + A_12
  A_1 ~ Alter_realW + Geschlecht 
  A_2 ~ Alter_realW + Geschlecht
  A_3 ~ Alter_realW + Geschlecht
  A_4 ~ Alter_realW + Geschlecht
  A_5 ~ Alter_realW + Geschlecht
  A_6 ~ Alter_realW + Geschlecht
  A_7 ~ Alter_realW + Geschlecht
  A_8 ~ Alter_realW + Geschlecht
  A_9 ~ Alter_realW + Geschlecht
  A_10 ~ Alter_realW + Geschlecht
  A_11 ~ Alter_realW + Geschlecht
  A_12 ~ Alter_realW + Geschlecht
"

modelG_p <- "
height =~ KGr_13 + KGr_14 + KGr_15 + KGr_16 + KGr_17 + KGr_18 + KGr_19 + KGr_20 + KGr_21 + KGr_22 + KGr_23 + KGr_24
  KGr_13 ~ Koerpergroesse_realW + Geschlecht
  KGr_14 ~ Koerpergroesse_realW + Geschlecht
  KGr_15 ~ Koerpergroesse_realW + Geschlecht
  KGr_16 ~ Koerpergroesse_realW + Geschlecht
  KGr_17 ~ Koerpergroesse_realW + Geschlecht
  KGr_18 ~ Koerpergroesse_realW + Geschlecht
  KGr_19 ~ Koerpergroesse_realW + Geschlecht
  KGr_20 ~ Koerpergroesse_realW + Geschlecht
  KGr_21 ~ Koerpergroesse_realW + Geschlecht
  KGr_22 ~ Koerpergroesse_realW + Geschlecht
  KGr_23 ~ Koerpergroesse_realW + Geschlecht
  KGr_24 ~ Koerpergroesse_realW + Geschlecht
"
modelGe_p <- "
  weight =~ KGe_25 + KGe_26 + KGe_27 + KGe_28 + KGe_29 + KGe_30 + KGe_31 + KGe_32 + KGe_33 + KGe_34 + KGe_35 + KGe_36
  KGe_25 ~ Koerpergewicht_realW + Geschlecht
  KGe_26 ~ Koerpergewicht_realW + Geschlecht
  KGe_27 ~ Koerpergewicht_realW + Geschlecht
  KGe_28 ~ Koerpergewicht_realW + Geschlecht
  KGe_29 ~ Koerpergewicht_realW + Geschlecht
  KGe_30 ~ Koerpergewicht_realW + Geschlecht
  KGe_31 ~ Koerpergewicht_realW + Geschlecht
  KGe_32 ~ Koerpergewicht_realW + Geschlecht
  KGe_33 ~ Koerpergewicht_realW + Geschlecht
  KGe_34 ~ Koerpergewicht_realW + Geschlecht
  KGe_35 ~ Koerpergewicht_realW + Geschlecht
  KGe_36 ~ Koerpergewicht_realW + Geschlecht
"

fit1A6_p <- cfa(model = modelA_p, data = dat_poly, ordered = c("A_1","A_2", "A_3", "A_4", "A_5", "A_6", "A_7","A_8", "A_9", "A_10", "A_11","A_12"), std.lv=TRUE, meanstructure=TRUE)
fit1A2_p <- cfa(model = modelA_p, data = dat_dich, ordered = c("A_1","A_2", "A_3", "A_4", "A_5", "A_6", "A_7","A_8", "A_9", "A_10", "A_11","A_12"), std.lv=TRUE, meanstructure=TRUE)
fit1G6_p <- cfa(model = modelG_p, data = dat_poly, ordered = c("KGr_13","KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19","KGr_20", "KGr_21", "KGr_22", "KGr_23","KGr_24"), std.lv=TRUE, meanstructure=TRUE)
fit1G2_p <- cfa(model = modelG_p, data = dat_dich, ordered = c("KGr_13","KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19","KGr_20", "KGr_21", "KGr_22", "KGr_23","KGr_24"), std.lv=TRUE, meanstructure=TRUE)
fit1Ge6_p <- cfa(model = modelGe_p, data = dat_poly, ordered = c("KGe_25","KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31","KGe_32", "KGe_33", "KGe_34", "KGe_35","KGe_36"), std.lv=TRUE, meanstructure=TRUE)
fit1Ge2_p <- cfa(model = modelGe_p, data = dat_dich, ordered = c("KGe_25","KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31","KGe_32", "KGe_33", "KGe_34", "KGe_35","KGe_36"), std.lv=TRUE, meanstructure=TRUE)

summary(fit1A6_p, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1A2_p, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1G6_p, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1G2_p, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1Ge6_p, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1Ge2_p, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
```

```{r}
bifac_df <- 
  bind_rows(
    standardizedSolution(fit1A6_p, type = "std.all") |> mutate(measure = "age", categories = "6"),
    standardizedSolution(fit1A2_p, type = "std.all") |> mutate(measure = "age", categories = "2"),
    standardizedSolution(fit1G6_p, type = "std.all") |> mutate(measure = "height", categories = "6"),
    standardizedSolution(fit1G2_p, type = "std.all") |> mutate(measure = "height", categories = "2"),
    standardizedSolution(fit1Ge6_p, type = "std.all") |> mutate(measure = "weight", categories = "6"),
    standardizedSolution(fit1Ge2_p, type = "std.all") |> mutate(measure = "weight", categories = "2")) |>
  filter(
    (lhs %in% c("age", "height", "weight") & op == "=~") |
    (rhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW", "Geschlecht") & op == "~")) |>
  mutate(
    cause = case_when(
      lhs %in% c("age", "height", "weight") & op == "=~" ~ lhs,
      rhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW", "Geschlecht") & op == "~" ~ rhs),
    item_name = case_when(
      lhs %in% c("age", "height", "weight") & op == "=~" ~ rhs,
      rhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW", "Geschlecht") & op == "~" ~ lhs
    )) |> 
  # Create new columns: prefix (everything before underscore) and suffix (the numeric part)
  mutate(prefix = sub("_.*", "", item_name),
         suffix = as.numeric(sub(".*_", "", item_name))) %>%
  # For each prefix group, assign new_num as the rank of suffix in sorted order 
  group_by(prefix) |>
  mutate(item = match(suffix, sort(unique(suffix)))) |>
  ungroup() |>
  mutate(est_print = papaja::printnum(est.std, gt1 = FALSE, digits = 3))
```

```{r}
table3 <- bifac_df |> 
  select(item, cause, measure, categories, est_print) |> 
  pivot_wider(names_from = c(measure, cause), values_from = est_print) |>
  flextable::as_grouped_data(groups = "categories") |>
  flextable::as_flextable() |>
  flextable::add_header_row(values = c("age", "height", "weight"),
    colwidths = c(4, 3, 3)) |>
  flextable::set_header_labels(
    values = list(
      item = "Item",
      age_age = "LV", height_height = "LV", weight_weight = "LV",
      age_Alter_realW = "age",
      height_Koerpergroesse_realW = "height",
      weight_Koerpergewicht_realW = "weight",
      age_Geschlecht = "sex", height_Geschlecht = "sex", weight_Geschlecht = "sex")) |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::border(i = 2, j = 1, border.top = officer::fp_border(width = 0), part = "header") |>
  flextable::border(i = 1, j = 1, border.bottom = officer::fp_border(width = 0), part = "header") |>
  flextable::theme_apa() |>
  ftExtra::colformat_md()
table3
```

### Full model

```{r}
modelges <- "
  age =~ A_1 +  A_2 + A_3 + A_4 + A_5 + A_6 + A_7 + A_8 + A_9 + A_10 + A_11 + A_12 
  height =~ KGr_13 + KGr_14 + KGr_15 + KGr_16 + KGr_17 + KGr_18 + KGr_19 + KGr_20 + KGr_21 + KGr_22 + KGr_23 + KGr_24 
  weight =~ KGe_25 + KGe_26 + KGe_27 + KGe_28 + KGe_29 + KGe_30 + KGe_31 + KGe_32 + KGe_33 + KGe_34 + KGe_35 + KGe_36
  Alter_realW ~~ Koerpergroesse_realW
  Koerpergewicht_realW ~~ Alter_realW
  Koerpergroesse_realW ~~ Koerpergewicht_realW

  Alter_realW ~~ age
  Koerpergewicht_realW ~~ weight
  Koerpergroesse_realW ~~ height
"

fit1ges6 <- cfa(model = modelges, data = dat_poly, ordered = c("A_1","A_2", "A_3", "A_4", "A_5", "A_6", "A_7","A_8", "A_9", "A_10", "A_11","A_12", "KGr_13","KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19","KGr_20", "KGr_21", "KGr_22", "KGr_23","KGr_24", "KGe_25","KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31","KGe_32", "KGe_33", "KGe_34", "KGe_35","KGe_36"), std.lv=TRUE, meanstructure=TRUE)
fit1ges2 <- cfa(model = modelges, data = dat_dich, ordered = c("A_1","A_2", "A_3", "A_4", "A_5", "A_6", "A_7","A_8", "A_9", "A_10", "A_11","A_12", "KGr_13","KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19","KGr_20", "KGr_21", "KGr_22", "KGr_23","KGr_24", "KGe_25","KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31","KGe_32", "KGe_33", "KGe_34", "KGe_35","KGe_36"), std.lv=TRUE, meanstructure=TRUE)

summary(fit1ges6, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
summary(fit1ges2, fit.measures=TRUE, standardized = TRUE, rsquare = TRUE, modindices = TRUE)
```

## Correlations between latent variables

```{r}
(r_latent_poly <- standardizedSolution(fit1ges6, type = "std.all") |>
  filter(op == "~~" & 
      lhs %in% c("age", "height", "weight") &
      rhs %in% c("age", "height", "weight") &
      lhs != rhs) |>
  mutate(est_print = papaja::printnum(est.std, gt1 = FALSE, digits = 3))) |>
  mutate(est_print = case_when(
    pvalue < 0.05 & pvalue > 0.01 ~ paste0(est_print, "*"),
    pvalue < 0.01 & pvalue > 0.001 ~ paste0(est_print, "**"),
    pvalue < 0.01 ~ paste0(est_print, "***"),
    .default = est_print 
  ))
r_latent_AG6 <- r_latent_poly |> filter(lhs == "age" & rhs == "height") |> pull(est_print)
r_latent_AGe6 <- r_latent_poly |> filter(lhs == "age" & rhs == "weight") |> pull(est_print)
r_latent_GGe6 <- r_latent_poly |> filter(lhs == "height" & rhs == "weight") |> pull(est_print)

(r_latent_dich <- standardizedSolution(fit1ges2, type = "std.all") |>
  filter(op == "~~" & 
      lhs %in% c("age", "height", "weight") &
      rhs %in% c("age", "height", "weight") &
      lhs != rhs) |>
  mutate(est_print = papaja::printnum(est.std, gt1 = FALSE, digits = 3))) |>
  mutate(est_print = case_when(
    pvalue < 0.05 & pvalue > 0.01 ~ paste0(est_print, "*"),
    pvalue < 0.01 & pvalue > 0.001 ~ paste0(est_print, "**"),
    pvalue < 0.01 ~ paste0(est_print, "***"),
    .default = est_print 
  ))
r_latent_AG2 <- r_latent_dich |> filter(lhs == "age" & rhs == "height") |> pull(est_print)
r_latent_AGe2 <- r_latent_dich |> filter(lhs == "age" & rhs == "weight") |> pull(est_print)
r_latent_GGe2 <- r_latent_dich |> filter(lhs == "height" & rhs == "weight") |> pull(est_print)
```

## Correlations between physical variables

```{r}
(r_real_poly <- corx(data = dat_poly |>
    select(Alter_realW, Koerpergroesse_realW, Koerpergewicht_realW),
  method = "pearson"))

r_real_AG6 <- r_real_poly$apa["Alter_realW", "Koerpergroesse_realW"]
r_real_AGe6 <- r_real_poly$apa["Alter_realW", "Koerpergewicht_realW"]
r_real_GGe6 <- r_real_poly$apa["Koerpergroesse_realW", "Koerpergewicht_realW"]

(r_real_dich <- corx(data = dat_dich |>
    select(Alter_realW, Koerpergroesse_realW, Koerpergewicht_realW),
  method = "pearson"))

r_real_AG2 <- r_real_dich$apa["Alter_realW", "Koerpergroesse_realW"]
r_real_AGe2 <- r_real_dich$apa["Alter_realW", "Koerpergewicht_realW"]
r_real_GGe2 <- r_real_dich$apa["Koerpergroesse_realW", "Koerpergewicht_realW"]
```

## Reliability estimates

```{r}
## ------------------------------------------------------------
## 1. recode inverse items on scales from 1-6 and 0-1
## ------------------------------------------------------------

# recode 6-point scale
recode_inverse_6 <- c('1' = 6L, '2' = 5L, '3' = 4L, '4' = 3L, '5' = 2L, '6' = 1L)

dat_poly_r <- dat_poly %>%
  mutate(across(all_of(c("A_10", "KGr_21", "KGr_22", "KGr_23", "KGe_31", "KGe_32", "KGe_33")),
                ~ recode(.x, !!!recode_inverse_6)))

# recode 2-point scale
recode_inverse_2 <- c('1' = 0L, '0' = 1L)

dat_dich_r <- dat_dich %>%
  mutate(across(all_of(c("A_10", "KGr_21", "KGr_22", "KGr_23", "KGe_31", "KGe_32", "KGe_33")),
                ~ recode(.x, !!!recode_inverse_2)))

## ------------------------------------------------------------
## 2. define item sets
## ------------------------------------------------------------

items_age <- c("A_1", "A_2", "A_3", "A_4", "A_5", "A_6", "A_7", "A_8", "A_9", "A_10", "A_11", "A_12")
items_height <- c("KGr_13", "KGr_14", "KGr_15", "KGr_16", "KGr_17", "KGr_18", "KGr_19", "KGr_20", "KGr_21", "KGr_22", "KGr_23", "KGr_24")
items_weight <- c("KGe_25", "KGe_26", "KGe_27", "KGe_28", "KGe_29", "KGe_30", "KGe_31", "KGe_32", "KGe_33", "KGe_34", "KGe_35", "KGe_36")

## ------------------------------------------------------------
## 3. Reliabilitätsanalyse (McDonald's Omega) mit Konfidenzintervallen
## ------------------------------------------------------------

# Für dat_poly (6-stufige Items)
(rel_A6 <- ci.reliability(data = dat_poly_r[, items_age], type = "omega", interval.type = "ml"))
rel_A6 <- papaja::printnum(rel_A6$est, gt1 = FALSE, digits = 3)
(rel_G6 <- ci.reliability(data = dat_poly_r[, items_height], type = "omega", interval.type = "ml"))
rel_G6 <- papaja::printnum(rel_G6$est, gt1 = FALSE, digits = 3)
(rel_Ge6 <- ci.reliability(data = dat_poly_r[, items_weight], type = "omega", interval.type = "ml"))
rel_Ge6 <- papaja::printnum(rel_Ge6$est, gt1 = FALSE, digits = 3)

# Für dat_dich (dichotome Items)
(rel_A2 <- ci.reliability(data = dat_dich_r[, items_age], type = "omega", interval.type = "ml"))
rel_A2 <- papaja::printnum(rel_A2$est, gt1 = FALSE, digits = 3)
(rel_G2 <- ci.reliability(data = dat_dich_r[, items_height], type = "omega", interval.type = "ml"))
rel_G2 <- papaja::printnum(rel_G2$est, gt1 = FALSE, digits = 3)
(rel_Ge2 <- ci.reliability(data = dat_dich_r[, items_weight], type = "omega", interval.type = "ml"))
rel_Ge2 <- papaja::printnum(rel_Ge2$est, gt1 = FALSE, digits = 3)
```

## Table 4

```{r}
cor_table <- tribble(
  ~Measure, ~Age_6     , ~Age_2     , ~Height_6   , ~Height_2   , ~Weight_6    , ~Weight_2    ,
  "Age"   , rel_A6     , rel_A2     , r_latent_AG6, r_latent_AG2, r_latent_AGe6, r_latent_AGe2,
  "Height", r_real_AG6 , r_real_AG2 , rel_G6      , rel_G2      , r_latent_GGe6, r_latent_GGe2,
  "Weight", r_real_AGe6, r_real_AGe2, r_real_GGe6 , r_real_GGe2 , rel_Ge6      , rel_Ge2      
)
```

```{r}
table4 <- flextable::flextable(cor_table) |>
  flextable::set_header_labels(
    Measure = "Categories",
    Age_6 = "6", Age_2 = "2",
    Height_6 = 6, Height_2 = 2,
    Weight_6 = 6, Weight_2 = 2) |>
  flextable::add_header_row(values = c("Measure", "Age", "Height", "Weight"),
    colwidths = c(1, 2, 2, 2)) |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::border(i = 2, j = 1, border.top = officer::fp_border(width = 0), part = "header") |>
  flextable::border(i = 1, j = 1, border.bottom = officer::fp_border(width = 0), part = "header") |>
  flextable::theme_apa() |>
  ftExtra::colformat_md()
table4
```

# Save Results for Manuscript

```{r}
results <- saveRDS(
  object = list(
    table3 = table3, table4 = table4
    ), 
  file = "../results/results.rds"
)
```