### 2025-06-30 analysis/tertiary_analyses.qmd
add efa analyses to supplement; format notbooks

---
title: "Tertiary analyses (scale construction)"
format:
  html:
    theme:
      - cosmo
      - ../notebooks.scss
    grid:
      body-width: 900px
---

::: {.callout-note appearance="simple"}

In this notebook, we reproduce the analyses of the scale construction dataset to select and reformulate the final item set used in the main study reported in the manuscript.

More details on the procedures that were used to construct the final item set used in the main study are reported in @eichhornItem2019, the German dissertation of the first author of the current manuscript.

:::

## Load packages

```{r}
library(foreign)
library(tidyverse)
library(psych)
```

```{r}
#| include: false
# record R packages
papaja::r_refs(file = "../r-references.bib")
```

## Load datasets

```{r}
dat_pilot <- read.spss("../data/Pilottest_Studie_3.sav", 
  to.data.frame = TRUE, use.value.labels = FALSE, na.omit = FALSE)
```

## Exploratory factor analysis

```{r}
fit <- fa(dat_pilot |> select(KGr_22:KGr_40), 
  nfactors = 3, rotate = "Promax", fm = "ml")

print(fit, sort = TRUE, cut = 0.15)
```

### Item total correlations

```{r}
neg_items <- c("A_13", "A_14", "A_21",
  "KGr_35", "KGr_36", "KGr_37", "KGr_38", "KGr_39", "KGr_40", "KGr_41", "KGr_43",
  "KGe_52", "KGe_53", "KGe_54", "KGe_55", "KGe_56", "KGe_57", "KGe_58")
dat_pilot_recode <- dat_pilot |> 
  select(KGr_22:KGr_40) |>
  mutate(across(all_of(neg_items), ~ 7 - .))

rel_age <- psych::alpha(dat_pilot_recode |> select(starts_with("A")))
rel_height <- psych::alpha(dat_pilot_recode |> select(starts_with("KGr")))
rel_weight <- psych::alpha(dat_pilot_recode |> select(starts_with("KGe")))
```

The part-whole corrected item total correlations are displayed in column `r.drop`

```{r}
rel_age$item.stats
rel_height$item.stats
rel_weight$item.stats
```

## References

::: {#refs}
:::